RILS-ROLS: Robust Symbolic Regression via Iterated Local Search and Ordinary Least Squares

### 2022-10-30 doc/paper.tex
Experimental evalaution: Section 4.4.

	 \item Concerning the high--accuracy of the \textsc{Rils}--\textsc{Rols} algorithm -- that is when termination of the algorithm yielded the error $R^2 \geq 0.999$ -- for the input data with the zero noise level on the small--sized random instance problems, it almost reaches the perfect score (i.e. $\approx$100\%). In the presence of a low noise level (equals to 0.001), the accuracy delivered by the algorithm is still high, at about 95\%. Just when a high level of noise (equals to 0.01) is presented in the input data, the accuracy stagnates, but still with the respectable 66\% small-sized instances it reaches at least a precision of $R^2=0.999$. However, as we said, a half of them math to the respectable exact models.

@@

	 \item Concerning the high--accuracy of the \textsc{Rils}--\textsc{Rols} algorithm on the large-sized instance problems,  the high--accuracy in the data free of the noise presence is reported for 60\% cases. Not so surprisingly, the hardest case to obtain the desired high--accuracy (that is, ensuring $R^2\geq 0.999$ upon termination) is in the presence of a high level of noise (of 0.01), where the algorithm was successful on about 20\% of the considered large--sized instance problems.

@@

Checking the  scalability of our \textsc{Rils}--\textsc{Rols} algorithm w.r.t.\ the number of variables presented in the corresponding exact models of the problem instances is shown by Figures~\ref{fig:compExact_noise_varcnt}--\ref{fig:compR2_noise_varcnt}. \fxnote{TODO}

\begin{center}
	\begin{tikzpicture}
		\begin{axis}[
			xlabel=$variable\;count$,
			ylabel=$exact\;percentage$,
			xmin=0, xmax=5,
			ymin=0, ymax=100,
			xtick={1,2,3,4},
			xticklabels={1,2,3,4},   % <---
			ytick={0,10,...,100}
			]
			\addplot[smooth,mark=*,blue] plot coordinates {
				(1,43.33)
				(2,43.33)
				(3,46.67)
				(4,46.67)
			};
			\addlegendentry{No noise}

@@

\begin{center}
	\begin{tikzpicture}
		\begin{axis}[
			xlabel=$variable\;count$,
			ylabel=$exact\;percentage$,
			xmin=0, xmax=5,
			ymin=0, ymax=100,
			xtick={1,2,3,4},
			xticklabels={1,2,3,4},   % <---
			ytick={0,10,...,100}
			]
			\addplot[smooth,mark=*,blue] plot coordinates {
				(1,83.33)
				(2,66.67)
				(3,73.33)
				(4,46.67)
			};
			\addlegendentry{No noise}

			\addplot[smooth,color=red,mark=x]
			plot coordinates {
				(1,56.67)
				(2,63.33)
				(3,66.67)
				(4,46.67)
			};
			\addlegendentry{Noise level 0.001}

			\addplot[smooth,color=green,mark=o]
			plot coordinates {
				(1,23.33)
				(2,43.33)
				(3,53.33)
				(4,26.67)
			};
			\addlegendentry{Noise level 0.01}
		\end{axis}
	\end{tikzpicture}
	\captionof{figure}{Percentages of solutions having $R^2 > 0.999$ for varying levels of noise and variable counts (only sizes ranging from 7 to 12)}
	\label{fig:compR2_noise_varcnt}
\end{center}

\subsection{Statistical evaluation}

\section{Conclusions and future work}\label{sec:conclusions}

>>>>

	 \item Concerning the high--accuracy of the \textsc{Rils}--\textsc{Rols} algorithm -- that is when termination of the algorithm yielded the error $R^2 \geq 0.999$ -- for the input data with the zero noise level on the small--sized random instance problems, it almost reaches the perfect score (i.e. $\approx$100\%). In the presence of a low noise level (equals to 0.001), the accuracy delivered by the algorithm is still high, at about 95\%. Just when a high level of noise (equals to 0.01) is presented in the input data, the accuracy stagnates, but still on the respectable 66\% small-sized instances it reaches a high--precision of $R^2=0.999$ upon termination. However, as we said, a half of them math to the respectable exact models.

@@

	 \item Concerning the high--accuracy of the \textsc{Rils}--\textsc{Rols} algorithm on the large-sized instance problems,  the high--accuracy in the data free of the noise presence is reported for 60\% cases. Not so surprisingly, the hardest case to obtain the desired high--accuracy (that is, ensuring n accuracy of $R^2\geq 0.999$ upon termination) is in the presence of a high level of noise (of 0.01), where the algorithm was successful on about 20\% of the considered large--sized instance problems.

@@

In this section we investigate the  scalability of our \textsc{Rils}--\textsc{Rols} algorithm w.r.t.\ the number of variables presented in the corresponding exact models of the problem instances whose sizes are ranging from 7 to 12. There is one line plot per each level of noise.  These are shown by Figure~\ref{fig:compExact_noise_varcnt}. The instances are grouped according to the different variable count that are included into respective exact models ($x$-axis). For each of the (four) group, the exact percentages of success rate of \textsc{Rils}--\textsc{Rols} algorithm is shown ($y$-axis). The following conclusions can be drawn from there.

The rate of success of \textsc{Rils}--\textsc{Rols} algorithm seems to slightly increase as the expected number of variables gets larger in case of the data with no noise;
	the increase is significant in the case of a higher noise in data. For example, the lowest percentages of success for \textsc{Rils}-\textsc{Rols} algorithm are noticed for noisy data where exact models consist of just one or two variables. It seems that in these cases the algorithm is more likely favorizing including additional variables at the cost of increasing model accuracy  in order to improve the fitness value over iterations. %Obtaining an accurate model with a few variables in the presence of a high level of nose in the data

\begin{center}
	\begin{tikzpicture}
		\begin{axis}[
			xlabel=$variable\;count\;in\;exact\;models$,
			ylabel=$exact\;percentage$,
			xmin=0, xmax=5,
			ymin=0, ymax=100,
			xtick={1,2,3,4},
			xticklabels={1,2,3,4},   % <---
			ytick={0,10,...,100}
			]
			\addplot[smooth,mark=*,blue] plot coordinates {
				(1,43.33)
				(2,43.33)
				(3,46.67)
				(4,46.67)
			};
			\addlegendentry{No noise}

@@

%\begin{center}
%	\begin{tikzpicture}
%		\begin{axis}[
%			xlabel=$variable\;count$,
%			ylabel=$exact\;percentage$,
%			xmin=0, xmax=5,
%			ymin=0, ymax=100,
%			xtick={1,2,3,4},
%			xticklabels={1,2,3,4},   % <---
%			ytick={0,10,...,100}
%			]
%			\addplot[smooth,mark=*,blue] plot coordinates {
%				(1,83.33)
%				(2,66.67)
%				(3,73.33)
%				(4,46.67)
%			};
%			\addlegendentry{No noise}
%			
%			\addplot[smooth,color=red,mark=x]
%			plot coordinates {
%				(1,56.67)
%				(2,63.33)
%				(3,66.67)
%				(4,46.67)
%			};
%			\addlegendentry{Noise level 0.001}
%			
%			\addplot[smooth,color=green,mark=o]
%			plot coordinates {
%				(1,23.33)
%				(2,43.33)
%				(3,53.33)
%				(4,26.67)
%			};
%			\addlegendentry{Noise level 0.01}
%		\end{axis}
%	\end{tikzpicture}
%	\captionof{figure}{Percentages of solutions having $R^2 > 0.999$ for varying levels of noise and variable counts (only sizes ranging from 7 to 12)}
%	\label{fig:compR2_noise_varcnt}
%\end{center}

\subsection{Statistical evaluation}
 \fxnote{TODO}
\section{Conclusions and future work}\label{sec:conclusions}
